ST563 Final Project

### 2021-04-23 DRAFT.Rmd
made the edits suggested by Dr. Griffith: Messed with the histograms and remade them in ggplot Added a references section and citations changed the part on page 4 where we talked about moderate, poor, and excellent classes. made the table rounded to 2 digits instead of a ton

There has been a push towards the implementation of different analytic solutions within the agriculture industry, especially in the realms of crop health and yield. These sort of analyses can help farmers reduce waste and improve profits. In particular, the viticulture industry stands to benefit from such methods, being that the production of wine is lengthy and multifaceted. Winemakers can work to optimize revenue by analyzing their input and output (crops and wines). Specifically, costs may be cut out if machine learning methods were able to predict the quality of wine based on different metrics, since wine producers would no longer have to hire expert wine tasters to determine wine quality. The main question the study looks to answer is can the quality of wine be predicted using a statistical model instead of taste testers.

@@

```{r}
# Create histograms for all 12 variables
par(mfrow = c(3,4))
for (i in 1:ncol(wine)){
  hist(wine[,i], xlab = names(wine)[i], 
       main = names(wine)[i])
}
```

Most wines are classified as moderate wines, with less wines falling in the excellent and poor categories. The density and pH of the wine seem to be normally distributed, while the sulfur, acid, and sulphate content appear skewed to the right. The density varies very little; looking at the scale on the distribution, one can see that it is approximately normally distributed around 0.997 with a standard deviation of about 0.002. The distribution of alcohol content came as no surprise; an average bottle of wine will range from about 10-12% alcohol by volume, with some exceptions. Beyond 16%, the wine is then considered fortified wine, and red wines don't usually have alcohol content below about 8%.

@@

kable(good.out, caption = "Summary Output for Good Wines") %>%
  kable_styling(latex_options = "HOLD_position")
kable(bad.out, caption = "Summary Output for Bad Wines") %>%
  kable_styling(latex_options = "HOLD_position")
```

@@

```{r prunedtree, out.height = '4in'}
# 2. pruned decision tree 
par(mfrow = c(1,2))
cv.tree.class <- cv.tree(tree.class, FUN = prune.misclass)
plot(cv.tree.class)
prune.tree.class <- prune.misclass(tree.class, best = 8)
plot(prune.tree.class); text(prune.tree.class, pretty = 0, cex = .6)
prune.tree.class.pred <- predict(prune.tree.class, tst, type = 'class')
prune.tree.class.train_err <- mean(trn$class != predict(prune.tree.class, type = 'class'))
prune.tree.class.test_err <- mean(tst$class != prune.tree.class.pred)
tree.res <- c(prune.tree.class.train_err, prune.tree.class.test_err)
# prune.tree.class.test_err
tree.res
```

@@

>>>>

There has been a push towards the implementation of different analytic solutions within the agriculture industry, especially in the realms of crop health and yield (Byrum et al. 2016). These sort of analyses can help farmers reduce waste and improve profits (Dimitriadis 2008). In particular, the viticulture industry stands to benefit from such methods, being that the production of wine is lengthy and multifaceted. Winemakers can work to optimize revenue by analyzing their input and output (crops and wines) (Santesteban 2019). Specifically, costs may be cut out if machine learning methods were able to predict the quality of wine based on different metrics, since wine producers would no longer have to hire expert wine tasters to determine wine quality. The main question the study looks to answer is can the quality of wine be predicted using a statistical model instead of taste testers.

@@

```{r}
# Create histograms for all 12 variables
hist1 <- ggplot(data = wine, aes(x = fixed.acidity)) + 
           geom_histogram(color = 'black', fill = 'grey', binwidth = 1) +
           labs(x = "Fixed Acidity", y = "Frequency")
hist2 <- ggplot(data = wine, aes(x = volatile.acidity)) + 
           geom_histogram(color = 'black', fill = 'grey', binwidth = .1) +
           labs(x = "Volatile Acidity", y = NULL)
hist3 <- ggplot(data = wine, aes(x = citric.acid)) + 
           geom_histogram(color = 'black', fill = 'grey', binwidth = .075) +
           labs(x = "Citric Acid", y = NULL)
hist4 <- ggplot(data = wine, aes(x = residual.sugar)) + 
           geom_histogram(color = 'black', fill = 'grey', binwidth = 1) +
           labs(x = "Residual Sugar", y = NULL)
hist5 <- ggplot(data = wine, aes(x = chlorides)) + 
           geom_histogram(color = 'black', fill = 'grey', binwidth = .05) +
           labs(x = "Chlorides", y = "Frequency")
hist6 <- ggplot(data = wine, aes(x = free.sulfur.dioxide)) + 
           geom_histogram(color = 'black', fill = 'grey', binwidth = 5) +
           labs(x = "Free Sulfur Dioxide", y = NULL)
hist7 <- ggplot(data = wine, aes(x = total.sulfur.dioxide)) + 
           geom_histogram(color = 'black', fill = 'grey', binwidth = 20) +
           labs(x = "Total Sulfur Dioxide", y = NULL)
hist8 <- ggplot(data = wine, aes(x = density)) + 
           geom_histogram(color = 'black', fill = 'grey', binwidth = .001) +
           labs(x = "Density", y = NULL)
hist9 <- ggplot(data = wine, aes(x = pH)) + 
           geom_histogram(color = 'black', fill = 'grey', binwidth = .1) +
           labs(x = "pH", y = "Frequency")
hist10 <- ggplot(data = wine, aes(x = sulphates)) + 
           geom_histogram(color = 'black', fill = 'grey', binwidth = .1) +
           labs(x = "Sulphates", y = NULL)
hist11 <- ggplot(data = wine, aes(x = alcohol)) + 
           geom_histogram(color = 'black', fill = 'grey', binwidth = .5) +
           labs(x = "Alcohol", y = NULL)
hist12 <- ggplot(data = wine, aes(x = factor(quality))) + 
           geom_bar(color = 'black', fill = 'grey') +
           labs(x = "Quality", y = NULL)
grid.arrange(hist1, hist2, hist3, hist4,
             hist5, hist6, hist7, hist8,
             hist9, hist10, hist11, hist12,
             nrow = 3, ncol = 4)
```

Most wines in the data set seem to be around the 5 to 7 range in terms of quality rating, meaning that there are not very many excellent or very poor wines. The density and pH of the wine seem to be normally distributed, while the sulfur, acid, and sulphate content appear skewed to the right. The density varies very little; looking at the scale on the distribution, one can see that it is approximately normally distributed around 0.997 with a standard deviation of about 0.002. The distribution of alcohol content came as no surprise; an average bottle of wine will range from about 10-12% alcohol by volume, with some exceptions. Beyond 16%, the wine is then considered fortified wine, and red wines don't usually have alcohol content below about 8%.

@@

kable(round(good.out, 2), caption = "Summary Output for Good Wines") %>%
  kable_styling(latex_options = "HOLD_position")
kable(round(bad.out, 2), caption = "Summary Output for Bad Wines") %>%
  kable_styling(latex_options = "HOLD_position")
```

@@

```{r prunedtree, out.height = '4in'}
# 2. pruned decision tree 
par(mfrow = c(1,2))
cv.tree.class <- cv.tree(tree.class, FUN = prune.misclass)
plot(cv.tree.class)
prune.tree.class <- prune.misclass(tree.class, best = 8)
plot(prune.tree.class); text(prune.tree.class, pretty = 0, cex = .6)
prune.tree.class.pred <- predict(prune.tree.class, tst, type = 'class')
prune.tree.class.train_err <- mean(trn$class != predict(prune.tree.class, type = 'class'))
prune.tree.class.test_err <- mean(tst$class != prune.tree.class.pred)
tree.res <- c(prune.tree.class.train_err, prune.tree.class.test_err)
# prune.tree.class.test_err
# tree.res
```

@@

# References

  * Byrum, Joseph, et al. "Advanced analytics for agricultural product development." Interfaces 46.1 (2016): 5-17.
  * Dimitriadis, Savvas, and Christos Goumopoulos. "Applying machine learning to extract new knowledge in precision agriculture applications." 2008 Panhellenic Conference on Informatics. IEEE, 2008.
  * Liakos, Konstantinos G., et al. "Machine learning in agriculture: A review." Sensors 18.8 (2018): 2674.
  * Santesteban, Luis G. "Precision viticulture and advanced analytics. A short review." Food chemistry 279 (2019): 58-62.
